Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-12 doc/manuscript/ESSDmanuscript.Rmd
work on C pools

@@

The end goal of IPCC greenhouse gas inventories is to quantify greenhouse gas emissions to, or withdrawals from, the atmosphere on an annual basis, most commonly on a national level [@ipcc_2006_2006;@ipcc_2019_2019]. 
For each stratum of subdivision within a land-use category, annual stock changes ($\Delta C$; Mg C yr^-1^) are calculated as the sum of changes in various pools, including any harvested wood products.
IPCC parses forest C pools into biomass (aboveground and belowground), dead organic matter (dead wood and litter), and soil organic matter (Table `r table_pools`).
For each pool, $\Delta C$ may be calculated using the "Gain-Loss Method", which takes the difference between gains and losses (influx and outflux variables in Fig. `r fig_variable_mapping`)<!--IPCC eq.2.4-->, or using the "Stock-Difference Method", which computes $\Delta C$ based on C stocks at two points in time<!--IPCC eq.2.5-->[@ipcc_2006_2006]. 
Thus, C cycle variables relevant to the IPCC methodology and to EFDB include C stocks, increments, and fluxes in the IPCC-defined pools (Fig. `r fig_variable_mapping`).

@@

The IPCC defines aboveground biomass as "all biomass of living vegetation, both woody and herbaceous, above the soil including stems, stumps, branches, bark, seeds, and foliage" [].

Belowground biomass is defined as "all biomass of live roots" [].

@@

Forest ecosystem C pools may be parsed in various ways, and while certain definitions and thresholds are more common than others, there is no single standard for measuring or reporting that is adhered to by all -- or even most -- studies.

>>>>

#Cited sections (listed here because these can change)
section_C_pools = 2.1

@@

The end goal of IPCC greenhouse gas inventories is to quantify greenhouse gas emissions to, or withdrawals from, the atmosphere on an annual basis, most commonly on a national level [@ipcc_2006_2006;@ipcc_2019_2019]. 
For each stratum of subdivision within a land-use category, annual stock changes ($\Delta C$; Mg C yr^-1^) are calculated as the sum of changes in various pools (section `r section_C_pools`), plus any harvested wood products.
For each pool, $\Delta C$ may be calculated using the "Gain-Loss Method", which takes the difference between gains and losses (influx and outflux variables in Fig. `r fig_variable_mapping`)<!--IPCC eq.2.4-->, or using the "Stock-Difference Method", which computes $\Delta C$ based on C stocks at two points in time<!--IPCC eq.2.5--> [@ipcc_2006_2006]. 
Thus, C cycle variables relevant to the IPCC methodology and to EFDB include C stocks, increments, and fluxes in the IPCC-defined pools (Fig. `r fig_variable_mapping`).

Forest ecosystem C pools may be parsed in various ways, and while certain definitions and thresholds are more common than others, there is no single standard for measuring or reporting that is adhered to by all -- or even most -- studies. 
IPCC parses forest C pools into biomass (aboveground and belowground), dead organic matter (dead wood and litter), and soil organic matter (Table `r table_pools`).
While there is some flexibility around the components included in each pool, each national inventory must apply these in a consistent manner.

@@

The IPCC defines aboveground biomass as "all biomass of living vegetation, both woody and herbaceous, above the soil including stems, stumps, branches, bark, seeds, and foliage" [@ipcc_2006_2006].

Belowground biomass is defined as "all biomass of live roots" [@ipcc_2006_2006].

@@

### 2022-04-12 doc/manuscript/ESSDmanuscript.tex
work on C pools

The end goal of IPCC greenhouse gas inventories is to quantify
greenhouse gas emissions to, or withdrawals from, the atmosphere on an
annual basis, most commonly on a national level
\citep{ipcc_2006_2006, ipcc_2019_2019}. For each stratum of subdivision
within a land-use category, annual stock changes (\(\Delta C\); Mg C
yr\textsuperscript{-1}) are calculated as the sum of changes in various
pools, including any harvested wood products. IPCC parses forest C pools
into biomass (aboveground and belowground), dead organic matter (dead
wood and litter), and soil organic matter (Table 1). For each pool,
\(\Delta C\) may be calculated using the ``Gain-Loss Method'', which
takes the difference between gains and losses (influx and outflux
variables in Fig. 1), or using the ``Stock-Difference Method'', which
computes \(\Delta C\) based on C stocks at two points in
time\citep{ipcc_2006_2006}. Thus, C cycle variables relevant to the IPCC
methodology and to EFDB include C stocks, increments, and fluxes in the
IPCC-defined pools (Fig. 1).

@@

\caption{\label{tab:table_pools}\textbf{IPCC-defined forest carbon pools with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{pool} & \textbf{definition} & \textbf{important sources of estimate variation} & \textbf{IPCC guidance}\\
\hline
aboveground biomass & all biomass of living vegetation, both woody and herbaceous & min dbh & acceptable to exclude understory if minor component\\
\hline
 &  & biomass allometry & \\
\hline
 &  & include dead standing? & no\\
\hline
belowground biomass & all biomass of live roots & min dbh (coarse roots) & \\
\hline
 &  & allometry or assumed ratio of belowground to aboveground biomass (IPCC table 4.4) & \\
\hline
 &  & min root diameter & fine roots may be excluded when they cannot be distinguished empirically from soil organic matter or litter\\
\hline
 &  & sampling depth & \\
\hline
dead wood & all non-living woody biomass not contained in the litter, either standing, lying on the ground, or in the soil & min diameter & 10 cm default, but may be chosen by country\\
\hline
 &  & include belowground? & \vphantom{1} yes\\
\hline
litter & all non-living biomass with a size greater than the limit for soil organic matter  and less than the minimum diameter chosen for dead wood, lying dead, in various states of decomposition above or within the mineral or organic soil & max diameter & 10 cm default, but may be chosen by country\\
\hline
 &  & layers included & includes entire O horizon\\
\hline
 &  & include belowground? & yes\\
\hline
soil organic matter & organic carbon in mineral soils to a specified depth & sampling depth & depth = 30 cm default, but may be chosen by country\\
\hline
\end{tabu}
\end{table}

@@

The IPCC defines aboveground biomass as ``all biomass of living
vegetation, both woody and herbaceous, above the soil including stems,
stumps, branches, bark, seeds, and foliage'' {[}{]}.

Belowground biomass is defined as ``all biomass of live roots'' {[}{]}.

>>>>

The end goal of IPCC greenhouse gas inventories is to quantify
greenhouse gas emissions to, or withdrawals from, the atmosphere on an
annual basis, most commonly on a national level
\citep{ipcc_2006_2006, ipcc_2019_2019}. For each stratum of subdivision
within a land-use category, annual stock changes (\(\Delta C\); Mg C
yr\textsuperscript{-1}) are calculated as the sum of changes in various
pools (section 2.1), plus any harvested wood products. For each pool,
\(\Delta C\) may be calculated using the ``Gain-Loss Method'', which
takes the difference between gains and losses (influx and outflux
variables in Fig. 1), or using the ``Stock-Difference Method'', which
computes \(\Delta C\) based on C stocks at two points in time
\citep{ipcc_2006_2006}. Thus, C cycle variables relevant to the IPCC
methodology and to EFDB include C stocks, increments, and fluxes in the
IPCC-defined pools (Fig. 1).

Forest ecosystem C pools may be parsed in various ways, and while
certain definitions and thresholds are more common than others, there is
no single standard for measuring or reporting that is adhered to by all
-- or even most -- studies. IPCC parses forest C pools into biomass
(aboveground and belowground), dead organic matter (dead wood and
litter), and soil organic matter (Table 1). While there is some
flexibility around the components included in each pool, each national
inventory must apply these in a consistent manner.

@@

\caption{\label{tab:table_pools}\textbf{IPCC-defined forest carbon pools with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{pool} & \textbf{definition} & \textbf{important sources of estimate variation} & \textbf{IPCC guidance}\\
\hline
aboveground biomass & all biomass of living vegetation & min dbh & acceptable to exclude understory if minor component\\
\hline
 &  & biomass allometry & \\
\hline
 &  & include dead standing? & no\\
\hline
belowground biomass & all biomass of live roots & min dbh (coarse roots) & \\
\hline
 &  & allometry or assumed ratio of belowground to aboveground biomass (IPCC table 4.4) & \\
\hline
 &  & min root diameter & fine roots may be excluded when they cannot be distinguished empirically from soil organic matter or litter\\
\hline
 &  & sampling depth & \\
\hline
dead wood & all non-living woody biomass not contained in the litter, either standing, lying on the ground, or in the soil & min diameter & 10 cm default, but may be chosen by country\\
\hline
 &  & include belowground? & \vphantom{1} yes\\
\hline
litter & all non-living biomass with a size greater than the limit for soil organic matter  and less than the minimum diameter chosen for dead wood, lying dead, in various states of decomposition above or within the mineral or organic soil & max diameter (= max diameter for deadwood) & 10 cm default, but may be chosen by country\\
\hline
 &  & layers included & includes entire O horizon\\
\hline
 &  & include belowground? & yes\\
\hline
soil organic matter & organic carbon in mineral soils to a specified depth & sampling depth & depth = 30 cm default, but may be chosen by country\\
\hline
\end{tabu}
\end{table}

@@

The IPCC defines aboveground biomass as ``all biomass of living
vegetation, both woody and herbaceous, above the soil including stems,
stumps, branches, bark, seeds, and foliage'' \citep{ipcc_2006_2006}.

Belowground biomass is defined as ``all biomass of live roots''
\citep{ipcc_2006_2006}.
